# Reproducible Research: Peer Assessment 1

```{r echo = TRUE,results = 'asis'}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
###1, Load the data
```{r Loading and preprocessing the data}
if(!file.exists('activity.csv')){
    unzip('repdata-data-activity.zip')
}
Data <- read.csv('activity.csv')
```
###2, Process/transform the data into a format suitable for your analysis
```{r echo = TRUE}
Data$date <- as.Date(Data$date, '%Y-%m-%d')
```

## what is mean total number of steps taken per day?
```{r echo = TRUE}
stepsByDay <- tapply(Data$steps, Data$date, sum, na.rm=TRUE)
str(stepsByDay)
```

###1, Make a histogram of the total number of steps taken each day
```{r echo = TRUE}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

###2, Calculate and report the mean and median total number of steps taken per day
```{r echo = TRUE}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```
The mean of steps taken per day is `r stepsByDayMean`; The median of steps taken per day is `r stepsByDayMedian`.

## What is the average daily activity pattern
```{r echo = TRUE}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=Data$steps), by=list(interval=Data$interval), FUN=mean, na.rm=TRUE)
str(averageStepsPerTimeBlock)
```

### 1, Make a time series plot
```{r echo = TRUE}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```
### 2, Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo = TRUE}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
head(mostSteps)
head(averageStepsPerTimeBlock[mostSteps,'interval'])
```

The maximum number of steps appear at time `r timeMostSteps`.

## Imputing missing values
###1, Calculate and report the total number of missing values in the dataset

```{r echo=TRUE}
numMissingValues <- sum(as.numeric(is.na(Data$steps)))
```
The total number of missing values is `r numMissingValues`.

###2, Devise a strategy for filling in all of the missing values in the dataset

###3, Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r echo =TRUE}
activityDataImputed <- Data
activityDataImputed$steps <- impute(Data$steps, fun=mean)
str(activityDataImputed)
```
###4, Make a histogram of the total number of steps taken each day
```{r echo =TRUE}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

### ... and Calculate and report the mean and median total number of steps taken per day.
```{r echo =TRUE}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
The mean of steps taken per day is `r stepsByDayMeanImputed`; The median of steps taken per day is `r stepsByDayMedianImputed`.

## Are there differences in activity patterns between weekdays and weekends?

###1, Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r echo = TRUE}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
str(activityDataImputed)
```
###2, Make a panel plot containing a time series plot
```{r echo = TRUE}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```